---
title: "Wheat Seed dataset"
author: "Molly Krzeczkowski"
date: "20/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      fig.retina = 3)
```

<!-- load packages  -->
```{r packages}
library(palmerpenguins)
library(tidyverse)
library(GGally)
library(ggplot2)
library(Rtsne)
```

# Introduction

The Wheat Seeds Dataset contains measurements of seed kernels from different varieties of wheat: Kama (1), Rosa (2) and Canadian (3). High quality visualization of the internal kernel structure was detected using a soft X-ray technique and 7 measurements were taken :Area, Perimeter, Compactness, Length of kernel, Width of kernel, Asymmetry coefficient, Length of kernel groove

We will investiage whether these variables might be useful to classify the species 

```{r data}
file <- "C:/Users/molsk/R work/Stage 3/week 8/Week-8-ws/raw_data/seeds_dataset.txt"
seed <- read.table(file, header = FALSE)
str(seed)
view(seed)
#rename columns
names(seed)[names(seed)=="V1"] <- "area"
names(seed)[names(seed)=="V2"] <- "perimeter"
names(seed)[names(seed)=="V3"] <-"compact"
names(seed)[names(seed)=="V4"] <- "length.of.kernel"
names(seed)[names(seed)=="V5"] <- "width_of_kernel"
names(seed)[names(seed)=="V6"] <- "asymmetry.coefficient"
names(seed)[names(seed)=="V7"] <- "length_of_kernel_groove"
names(seed)[names(seed)=="V8"] <- "strain"

view(seed)
```

<!-- clean the variable names -->
```{r clean}
wheat <- seed %>%
  janitor::clean_names()
# look at the 8 variables 
wheat %>% select(everything()) %>% 
  summary()

# strain is numerical lets make it catagorical 
wheat$strain <- recode(wheat$strain,
                       '1' = "Karma",
                       '2' = "Rosa",
                       '3' = "Canadian")
view(wheat)
```
<!-- No values are missing  -->
```{r visualise}
wheat %>% 
  select(everything()) %>% 
  ggpairs(aes(color = strain)) +
  scale_color_manual(values = c("purple", "red", "yellow"))
```
<!-- look at pca -->
```{r pca}
pca <- wheat %>% 
  select(-strain) %>%
  prcomp(scale. = TRUE)

# view pca
summary(pca)
# nearly 99% varaition is in the first 3 
pca$rotation

# extract the scores into a dataframe with the strain names
pca_labelled <- data.frame(pca$x, strain = wheat$strain)

# make a scatterplot 
pca_labelled %>% ggplot(aes(x = PC1, y = PC2, color = strain)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("purple", "red", "yellow"))
```
<!-- t-SNE is a non-parametric, non-linear method that prioritises placing similar observations near each other. -->
<!-- It is a probabilistic method and is computational intensive. -->
```{r tsne}
tsne <- wheat %>% 
  select(-strain) %>% 
  Rtsne(perplexity = 20, check_duplicates = FALSE)

datwheat <- data.frame(tsne$Y, strain = wheat$strain)
datwheat %>% ggplot(aes(x = X1, y = X2, colour = strain)) + geom_point()
```

