---
title: "Wheat Seed dataset"
author: "Molly Krzeczkowski"
date: "20/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      fig.retina = 3)
```

<!-- load packages  -->
```{r packages}
library(palmerpenguins)
library(tidyverse)
library(GGally)
library(ggplot2)
```

# Introduction

The Wheat Seeds Dataset contains measurements of seed kernels from different varieties of wheat: Kama (1), Rosa (2) and Canadian (3). High quality visualization of the internal kernel structure was detected using a soft X-ray technique and 7 measurements were taken :Area, Perimeter, Compactness, Length of kernel, Width of kernel, Asymmetry coefficient, Length of kernel groove

We will investiage whether these variables might be useful to classify the species 

```{r data}
file <- "C:/Users/molsk/R work/Stage 3/week 8/Week-8-ws/raw_data/seeds_dataset.txt"
seed <- read.table(file, header = FALSE)
str(seed)
view(seed)
#rename columns
names(seed)[names(seed)=="V1"] <- "area"
names(seed)[names(seed)=="V2"] <- "perimeter"
names(seed)[names(seed)=="V3"] <-"compact"
names(seed)[names(seed)=="V4"] <- "length.of.kernel"
names(seed)[names(seed)=="V5"] <- "width_of_kernel"
names(seed)[names(seed)=="V6"] <- "asymmetry.coefficient"
names(seed)[names(seed)=="V7"] <- "length_of_kernel_groove"
names(seed)[names(seed)=="V8"] <- "strain"

view(seed)
```

